---
output: github_document
editor_options: 
  chunk_output_type: console
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, echo = FALSE}
knitr::knit_hooks$set(margin = function(before, options, envir) {
  if (before) par(mgp = c(1.5, .5, 0), bty = "n", plt = c(.105, .97, .13, .97))
  else NULL
})

knitr::opts_chunk$set(margin = TRUE, prompt = TRUE, comment = "#>",
                      collapse = TRUE, cache = FALSE, autodep = TRUE,
                      dev.args = list(pointsize = 11), fig.height = 3.5,
                      fig.width = 4.24725, fig.retina = 2,
                      fig.align = "center", fig.path = "README-")
```

# gamar

The package `gamar` provides an interface to the GAMA simulation platform 
([gama-platform.org](gama-platform.org)). It loads gaml files defining models
and provides facilities to define plans of experiments. It also call the GAMA
engine to run defined plans of experiments. The simulation results can also be
loaded into R for subsequent analyses.

## Installation

Before installing `gamar` you need to install the package `devtools` if not
already installed on your system:

```{r eval = FALSE}
if (!"devtools" %in% rownames(installed.packages())) install.packages("devtools")
```

Once `devtools` is installed on your system, you can install `gamar` directly
from GitHub:

```{r eval = FALSE}
devtools::install_github("choisy/gamar")
```


## Usage

The package `gamar` contains 6 functions:

* `defpath` to define the path to the GAMA executable
* `examples` to manipulate examples from the built-in library
* `list_gaml` to list the paths to the gaml models files in a hierarchy
* `list_experiment` to list experiments of a gaml model file
* `experiment` to create an object of class `experiment`
* `load_experiment` to load an experiment of a model

See articles and vignettes for more details.
