---
title: "Using gamar"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r setup, include = FALSE}
knitr::knit_hooks$set(margin = function(before, options, envir) {
  if (before) par(mgp = c(1.5, .5, 0), bty = "n", plt = c(.105, .97, .13, .97))
  else NULL
})

knitr::opts_chunk$set(margin = TRUE, prompt = TRUE, comment = "",
                      collapse = TRUE, cache = FALSE, autodep = TRUE,
                      dev.args = list(pointsize = 11), fig.height = 3.5,
                      fig.width = 4.24725, fig.retina = 2, fig.align = "center")
```

```{r include = FALSE}
workspace <- "/Users/choisy/using_gamar"
if (dir.exists(workspace)) unlink(workspace, recursive = TRUE)
dir.create(workspace)
knitr::opts_knit$set(root.dir = workspace)
rm(list = ls())
```



## Introduction

To be able to use `gamar` you need to install GAMA (> 1.7) on your system. You
can do it from [gama-platform.org](gama-platform.org). You then need to
install the `gamar` R package, load it and configure it before you can use it.
The `gamar` package contains 6 functions:

* `defpath` to define the path to the GAMA executable
* `examples` to manipulate examples from the built-in library
* `get_gaml` to list the paths to the gaml models files in a hierarchy
* `get_experiment` to list experiments of a gaml model file
* `experiment` to create an object of class `experiment`
* `load_experiment` to load an experiment of a model

## Installing gamar

Before installing `gamar` you need to install the package `devtools` if not
already installed on your system:

```{r eval = FALSE}
if (!"devtools" %in% rownames(installed.packages())) install.packages("devtools")
```

Once `devtools` is installed on your system, you can install `gamar` directly
from GitHub:

```{r eval = FALSE}
devtools::install_github("choisy/gamar")
```


## Loading and configuring gamar

With `gamar` installed on your system, you need to load it before using it:

```{r}
library(gamar)
```

If you are using OS X or macOS, `gamar` will by default use the
`/Applications/Gama.app` if it exists. If not, you can manually define the path
to the GAMA with the `defpath` function:

```{r eval = FALSE}
defpath("path_to_your_GAMA_folder")
```


## Built-in models library


There are a number of built-in examples in the `gamar` package, of which you
can see the list:

```{r}
examples()
```

To use one of these examples, we recommand that you first move it to your local
directory (by default):

```{r}
dir()
examples("sir")
dir()
```

or to any other directory of your choice:

```{r}
examples("sir", "one_directory")
dir()
dir("one_directory")
```

If a file with the same name already exists in the destination directory, the
name of the copied file is appended a digit making it unique:

```{r}
examples("sir", "one_directory")
dir("one_directory")
```

The minimum that these examples include is a `.gaml` file that specifies the
model and some experiment(s). Sometimes additional files are also included, for
examples, shape files defining the space. That's the case for the `predator-prey`
model. All the files of the model are then contained in a directory:

```{r}
dir()
examples("predator-prey")
dir("predator_prey", recursive = TRUE)
```

Again, if the name of the directory already exists in the destination directory,
the name of the directory is appended a digit making it unique:

```{r}
examples("predator-prey")
dir()
```

Last thing about this `examples` function is that you can copied several models
at the same time:

```{r}
examples(c("sir", "predator-prey"))
```

specifying a common destination directory:

```{r}
examples(c("sir", "predator-prey"), "common_directory")
dir()
dir("common_directory")
```

or a different directory for each model:

```{r}
examples(c("sir", "predator-prey"), c("directory1", "directory2"))
dir()
dir("directory1")
dir("directory2")
```


## Loading experiments

Listing the GAMA models in a file hierarchy:

```{r}
(mod <- list_gaml("predator_prey"))
```

The argument to `get_gaml` is the path to the top of a file hierarchy. By
default, it's the current directory. Listing the experiments of a GAMA model:

```{r}
list_experiment(mod)
```

Loading a model from an experiment:

```{r}
predprey <- load_experiment("prey_predator", mod)  # creates workgamar with sir_1.xml in it
```

Similarly with the SIR model:

```{r}
dir()
list_experiment("sir.gaml")
sir_exp <- load_experiment("sir", "sir.gaml")
```

**TO DO: loading several experiments from a same model GAMA file.**

```{r}
class(sir_exp)
```

`print` method:

```{r}
sir_exp
```

Underlying `list` structure:

```{r}
is.list(sir_exp)
str(sir_exp)
names(sir_exp)
```



## Manipulating experiments

Values of an experiment can be changed as for any list

```{r}
sir_exp$parameters["I0"]
sir_exp$parameters["I0"] <- 10
sir_exp$parameters["I0"]
```

Editing the model:

```{r eval = FALSE}
edit_model(sir_exp)
```


